%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% chapter4.tex
%% NOVA thesis document file
%%
%% Chapter with lots of dummy text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE chapter4.tex}%

\chapter{Introduction}
\label{cha:introduction}

Software architecture is one of the most important factor's that influences the capabilities and constraints of a system's evolution throughout its lifecycle \cite{Breivold2012}.
The subject of architecture is primarily concerned with the composition and design of the fundamental structures of a software system,
with the goal of meeting functional requirements, while maximizing non-functional criteria, such as scalability, maintainability, extensibility, availability, and so on.

\paragraph{}

A software architectural model may be viewed as a macroscopic blueprint for the composition of a system elements.
Element's typically fall into one of three overlapping categories: data elements, processing elements, and connecting elements. \cite{architecture}.
The data elements contain the state of system (eg. database, variable, file);
The processing elements transform the data elements (eg. method, function, service);
The connecting elements serve as the bridge that permits different processing elements to collaborate (e.g. procedure call, shared memory, network message).
Software architectural models are primarily distinguished by the capabilities of the connecting elements and by the layers used to segregate elements into different subsystems.
This thesis aims to enhance the connecting elements used in microservice architectures in order to fulfill stricter extensibility and availability requirements.

\paragraph{}

The field of software architecture traces back to 1960s, but it wasn't until 1992 that \citeauthor{architecture} \cite{architecture} created a firm basis on the subject.
Changing the architecture of a system after its initial design is typically very costly.
Software engineers have developed novel methods for composing systems that offer broad functionality and fit a diverse set of criteria, \citeauthor{3} work \cite{3, 4} offers an overview of software architectural research.

\paragraph{}

A typical example of an architectural design is the Model-View-controller (MVC) \cite{mvc, microservices}.
The MVC pattern encourages three separate layers: The (M)odel layer manages the data of an application;
The (V)iew layer handles the display of information to an end user; The (C)ontroller layer
processes user input and converts it to commands for the model or view layers.
These layers improve software maintainability by separating distinct concerns,
however this architectural model is notorious for its lack of extensibility and scalability due to monolithic nature of each layer.
Creating and modifying instances of massive monoliths is time-consuming and error-prone;
minor system updates require the complete redeployment of the application, severely limiting a system's capacity to evolve while preserving its availability.

\paragraph{}

Service Oriented Architecture (SOA) \cite{7} is an alternative model that aims to
improve the re-usability and maintainability of software by minimizing the coupling between independent system components.
In SOA, components provide functionality to other components via network message passing.
A component is exposed via the service abstraction, which defines a standardized contract for the message exchange and format.
Microservices \cite{microservices} are a modern iteration of the SOA model,
that reduces the excessive and imposing architectural layers introduced in earlier generations of the SOA model.

\paragraph{}

Some advantages of service orientation over monoliths are as follows:
\begin{itemize}
    \item Functional independence - Each component is operationally independent, components only communicate through their standardized interfaces.
    \item Component scalability - When a single component becomes overloaded, additional instances of the component can be deployed separately.
    \item Separate development - Different teams may work on different modules of the system with minimal coordination if the component interfaces are defined ahead of time.
    \item Flexibility - Each component can be implemented in the programming language and platform that best suits its requirements.
\end{itemize}

These advantages boost the system's ability to evolve by allowing orthogonal domain's to be disassembled into disjoint subsystems.
In a microservice architecture, subsystems can fully evolve independently of one another, as long as they don't share services.
Because services are tied to one another via their interfaces, they can only evolve separately if their contracts remain consistent.
It is often advantageous to have the capacity to change service contracts independently,
particularly in agile workflows and the early phases of development, where contracts are prone to frequent changes.

\section{Problem Statement} % (fold)
\label{sec:problem_statement}

The main challenge in developing applications based on microservices-based architectures is the lack of mechanisms for evaluating the safety of service contract updates.

\paragraph{}

In a monolithic system, interactions between different components are conducted through procedure calls.
In a distributed system, however, each individual component has to communicate with other components across the network,
without the same guarantees.

\paragraph{}

Modifying a method definition and all references in a monolithic application may be done quickly with the aid of IDE refactoring tools.
Equivalent tools for detecting incompatibilities between endpoints and their consumers are scarce in distributed systems.
Developers are left with the burden of manually tracking down and refactoring the system's dependencies across all consumers and producers services.

\paragraph{}

With a rising number of separate services and their interactions, contract administration and service integration become progressively more challenging.
Software engineering guidelines assume a scenario of frequent service implementation changes and few contract modifications;
in the context of iterative approaches to software development, contracts are typically modified as frequently as implementation details.

\paragraph{}

Preserving the integrity of microservice architectures is a daunting task that can only be accomplished by the most diligent teams, for the following reasons:

\paragraph{ The ramifications of changing a contract are discovered after the fact:}
Services evolve independently, and contracts are rarely formally specified or documented.
The only symptoms of a broken system are runtime errors or unexpected behaviors.

\paragraph{ The safety of a deployment operation is unknown:}
Established orchestration platforms can't guarantee the security of deployments because they rely only on software packaging information (like version IDs) to manage deployments.
A new model for such environments is required to ensure the safety of deployment tasks, one that uses information from contracts of previously deployed services,
 and employs a compatibility relation between them.

\paragraph{ A contract change, entails the redeployment and downtime of its consumers:}
Modifications to a service's definition have an immediate impact on the entire system, by requiring the downtime of all consumer services.
Deployments of “compatible” services should not disrupt the system soundness,
one service should be able to be replaced while the remainder of the system remains operational.

\section{Contributions} % (fold)
\label{sec:contributions}

The current  work  aims to promote contract evolution safety and to ensure that changes in producers contracts do not require the consumers upgrade and redeployment.

\paragraph{}

The challenge of insuring the safety of contract evolutions will be handled by defining and adopting an API description language, as well as implementing
a pre-flight safety check procedure that verifies whether service contracts are compatible before deployment.
The compatibility mechanism is akin to type-checking procedures in compiled languages.
The aforementioned mechanism preserves a distributed system's correct behavior by rejecting additions or modifications that would threaten it and by informing the programmer in advance
of all the repercussions of the deployment.

\paragraph{}

To ensure that changes in producers contracts do not require the consumers upgrade and redeployment, contract evolution
will be supported at runtime by a generated proxy component that dynamically adapts the data exchanged between services.
The proxy adapter is capable of evaluating all the outgoing and incoming requests and determine whether
they should be left unchanged or how they should be transformed.

\paragraph{}

The evolution specification between compatible contracts is semi-automated and human-readable.
The addition, deletion, rename and migration of data fields in a producer modules will be supported under the proposed evolution specification language and system.
Modifications on the signatures of WebAPIs, such as changing a path parameter to a query parameter, will also be supported.
More complex changes in data fields, such as type or format changes,
will be supported through the use of user-defined functions and a language interpreter.
In summary, this work will provide:

\begin{itemize}
    \item A compatibility verification on service contracts that determines whether or not messages may be exchanged without data loss;
    \item Definition of a core language for the specification of contract compatibility corrections;
    \item Implementation of a robust type-safe adaptation protocol for service contracts;
    \item Implementation of a management system tasked with holding service contracts, tracking service dependencies,
    and providing a lightweight preflight safety check procedure for deployment/un-deployment operations
    \item A benchmark platform to evaluate the solution.
\end{itemize}

\section{Document Scructure} % (fold)
\label{sec:document_structure}

The document is organized into seven sections, which are: Chapter 1-Introduciton, Chapter 2-
Background, Chapter 3 - Related work; Chapter 4-System design, Chapter 5 - Implementation;
Chapter 6- Results and Chapter 7- Conclusions and future work.

\paragraph{}

The first chapter, Introduction, discusses the source of the problem as well as its challenges,
and the proposed work's contributions.

\paragraph{}

The second chapter, Background, is a review of the main concepts, techniques, and
applications underlying the problem, and the proposed approach.
The ideas are expanded upon in greater detail in the third chapter, which evaluates the relevant work on topics such as
Web APIs, schema representation languages, schema registry, service e contract evolution,
service adapters, deployment Strategies, and API Management Tools.

\paragraph{}

In the fourth chapter, System Design, the design of the proposed system is presented, as well
as some alternatives and decisions made.
The first section of this chapter is devoted to the
criteria that guided the system's design, while the remainder of the chapter is devoted to a
discussion of the considered relevant aspect.

\paragraph{}

The implementation of the proposed system is presented in the fifth chapter, Implementation.
It begins with an overview of the system's architecture before delving deeply into the implementation of each module.

\paragraph{}

The main results of the implementation are assessed in the sixth chapter, Results.
The benchmark platform is presented in this chapter, as well as the methodology and experiments carried out.

\paragraph{}

Finally, the last chapter is dedicated to the main conclusions and recommendations of future
work.

